Windowed least-squares model reduction for dynamical systems

### 2019-10-23 section3/general_form.tex
active voice

%\subsection{Formulation of the Time-Continuous LSPG Method} In contrast to
%the LSPG approach, which minimizes the fully discrete residual over a
%time-step, the present work proposes minimizing the time-continuous FOM ODE
%residual over time slabs. The proposed method can be formulated equivalently
%from two viewpoints. The first approach formulates the problem from a
%standard calculus of variations viewpoint and is analogous to the classic
%residual minimization process used in the LSPG method (but here performed at
%the continuous level). The second viewpoint corresponds to formulating an
%optimal control problem in Lagrange form (which is a specific instance of an
%optimal control problem of Bolza type). In this viewpoint, the
%\methodAcronym\ is formulated as an optimal control problem in where the
%objective is to find a ``controller" that minimizes the residual of the
%Galerkin ROM. We now outline both of these (equivalent) formulations. 
%
\subsection{Windowed least-squares for general space--time trial subspaces} 
We begin by introducing a (potentially nonuniform) partition of the time domain $[0,T]$
into $\nslabs$ non-overlapping windows $[\timeStartArg{n} ,
\timeEndArg{n}]\subseteq[0,T]$ of length $\DeltaSlabArg{n}\defeq\timeEndArg{n} -
\timeStartArg{n}$, $n=1,\ldots,\nslabs$ such that 
$\timeStartArg{1} = 0$, $\timeEndArg{\nslabs} = T$, and
$\timeStartArg{n+1} = \timeEndArg{n}$,
$n=1,\ldots,\nslabs-1$; Fig.~\ref{fig:slab_fig} depicts such a partitioning.
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} 
\caption{Partitioning of the time
	domain into time windows.} 
\label{fig:slab_fig} 
\end{centering} 
\end{figure}
Over the $n$th time window, we approximate the FOM ODE solution as 
$\approxstate^n(t)\approx \stateFOM(t)$,
$t\in[\timeStartArg{n},\timeEndArg{n}]$, which is enforced to reside in 
the $n$th space--time trial subspace 
$\stspaceArg{n}$
such that
\begin{equation}
\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,
\end{equation}
where  $\timeSpaceArg{n}$ denotes the set of (sufficiently smooth) real-valued functions acting on
$[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} =
\{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$).
%and
%$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. 
For notational purposes, we additionally define the (spatial) trial subspaces at the start of each window as
\begin{equation}
	\stspaceBoundStartArg{n}\defeq\mathrm{span}(\{\statey(\timeStartArg{n})\,|\, \statey\in
	\stspaceArg{n}\})\subseteq\RR{\fomdim}, \qquad n=1,\ldots,\nslabs, 
\end{equation}
such that $\approxstateArg{n}{\timeStartArg{n}} \in \stspaceBoundStartArg{n}$.
To outline \methodAcronym, we now define the objective functional over the
$n$th window as
\begin{equation}\label{eq:obj}
\begin{split} \mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArgt{n}{t} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t, \\
%&\vcentcolon \RR{\fomdim} \times [\timeStartArg{n},\timeEndArg{n}] \rightarrow
%\RR{}, \end{align}
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
\RRplus, 
\end{split}
\end{equation}
%\begin{align}\label{eq:obj} \mathcal{J}^n &\vcentcolon \state \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \resid(\state)^T
%\stweightingMat \resid(\state) d\tau \\ &\vcentcolon \RR{\fomdim} \times
%[0,T] \rightarrow \RR{}.  \end{align}
where $\stweightingMatArg{n} \equiv \stweightingMatOneTArg{n}
\stweightingMatOneArg{n} \in \RR{\fomdim \times \fomdim}$ denotes a
symmetric positive semi-definite matrix that can enable hyper-reduction, for
example. 
%For notational simplicity, we employ the same weighting matrix for all time windows.

The \methodAcronym\ framework sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
$\approxstateArgnt{n}$ is the solution to the
minimization problem
\begin{equation}\label{eq:tclsrm}
\begin{split}
      &\underset{\statey \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\statey), \\
			&\text{subject to } \;  \statey(\timeStartArg{n}) =
\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1}(\timeEndArg{n-1})) & n = 2,\ldots,\nslabs, \\
\mathbb{P}^n( \stateFOMIC)& n=1, \end{cases} 
\end{split}
\end{equation}
%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
%\begin{equation}\label{eq:tclsrm}
%\begin{split}
%       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
%      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
%\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
%\end{split}
%\end{equation}
where $\projectorArg{n} : \RR{\fomdim} \rightarrow \stspaceBoundStartArg{n}$ 
is a (spatial) projection operator onto the start of the $n$th trial space
(e.g., $\elltwo$-orthogonal projection operator).

>>>>

%\subsection{Formulation of the Time-Continuous LSPG Method} In contrast to
%the LSPG approach, which minimizes the fully discrete residual over a
%time-step, the present work proposes minimizing the time-continuous FOM ODE
%residual over time slabs. The proposed method can be formulated equivalently
%from two viewpoints. The first approach formulates the problem from a
%standard calculus of variations viewpoint and is analogous to the classic
%residual minimization process used in the LSPG method (but here performed at
%the continuous level). The second viewpoint corresponds to formulating an
%optimal control problem in Lagrange form (which is a specific instance of an
%optimal control problem of Bolza type). In this viewpoint, the
%\methodAcronym\ is formulated as an optimal control problem in where the
%objective is to find a ``controller" that minimizes the residual of the
%Galerkin ROM. We now outline both of these (equivalent) formulations. 
%
\subsection{Windowed least-squares for general space--time trial subspaces} 
We begin by introducing a (potentially nonuniform) partition of the time domain $[0,T]$
into $\nslabs$ non-overlapping windows $[\timeStartArg{n} ,
\timeEndArg{n}]\subseteq[0,T]$ of length $\DeltaSlabArg{n}\defeq\timeEndArg{n} -
\timeStartArg{n}$, $n=1,\ldots,\nslabs$ such that 
$\timeStartArg{1} = 0$, $\timeEndArg{\nslabs} = T$, and
$\timeStartArg{n+1} = \timeEndArg{n}$,
$n=1,\ldots,\nslabs-1$; Fig.~\ref{fig:slab_fig} depicts such a partitioning.
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} 
\caption{Partitioning of the time
	domain into time windows.} 
\label{fig:slab_fig} 
\end{centering} 
\end{figure}
Over the $n$th time window, we approximate the FOM ODE solution as 
$\approxstate^n(t)\approx \stateFOM(t)$,
$t\in[\timeStartArg{n},\timeEndArg{n}]$, which is enforced to reside in 
the $n$th space--time trial subspace 
$\stspaceArg{n}$
such that
\begin{equation}
\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,
\end{equation}
where  $\timeSpaceArg{n}$ denotes the set of (sufficiently smooth) real-valued functions acting on
$[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} =
\{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$).
%and
%$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. 
For notational purposes, we additionally define the spatial trial subspaces at the start of each window as
\begin{equation}
	\stspaceBoundStartArg{n}\defeq\mathrm{span}(\{\statey(\timeStartArg{n})\,|\, \statey\in
	\stspaceArg{n}\})\subseteq\RR{\fomdim}, \qquad n=1,\ldots,\nslabs, 
\end{equation}
such that $\approxstateArg{n}{\timeStartArg{n}} \in \stspaceBoundStartArg{n}$.
To outline \methodAcronym, we now define the objective functional over the
$n$th window as
\begin{equation}\label{eq:obj}
\begin{split} \mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArgt{n}{t} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t, \\
%&\vcentcolon \RR{\fomdim} \times [\timeStartArg{n},\timeEndArg{n}] \rightarrow
%\RR{}, \end{align}
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
\RRplus, 
\end{split}
\end{equation}
%\begin{align}\label{eq:obj} \mathcal{J}^n &\vcentcolon \state \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \resid(\state)^T
%\stweightingMat \resid(\state) d\tau \\ &\vcentcolon \RR{\fomdim} \times
%[0,T] \rightarrow \RR{}.  \end{align}
where $\stweightingMatArg{n} \equiv \stweightingMatOneTArg{n}
\stweightingMatOneArg{n} \in \RR{\fomdim \times \fomdim}$ denotes a
symmetric positive semi-definite matrix that can enable hyper-reduction, for
example. 
%For notational simplicity, we employ the same weighting matrix for all time windows.

The \methodAcronym\ \approachKwd\ sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
$\approxstateArgnt{n}$ is the solution to the
minimization problem
\begin{equation}\label{eq:tclsrm}
\begin{split}
      &\underset{\statey \in \stspace^n}{\text{minimize}} \; \mathcal{J}^n(\statey), \\
			&\text{subject to } \;  \statey(\timeStartArg{n}) =
\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1}(\timeEndArg{n-1})) & n = 2,\ldots,\nslabs, \\
\mathbb{P}^n( \stateFOMIC)& n=1, \end{cases} 
\end{split}
\end{equation}
%\methodAcronym\ computes a sequence of approximate solutions $\approxstateArgnt{n}$ as the solution to the minimization problem,
%\begin{equation}\label{eq:tclsrm}
%\begin{split}
%       &\approxstateArgnt{n} =  \underset{\statey \in \stspace^n}{\text{arg\,min}} \; \mathcal{J}^n(\statey), \\ 
%      &\text{subject to }\;  \approxstateArg{n}{\timeStartArg{n}} =
%\begin{cases}\mathbb{P}^n (\approxstateArgnt{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs \\
%\mathbb{P}^n( \stateFOMArgnt{1})(0)& n=1, \end{cases} 
%\end{split}
%\end{equation}
where $\projectorArg{n} : \RR{\fomdim} \rightarrow \stspaceBoundStartArg{n}$ 
is a (spatial) projection operator onto the start of the $n$th trial space
(e.g., $\elltwo$-orthogonal projection operator).
